---
title: "p8105_hw1_lah2227"
author: "Lauren Holley"
date: "2026-09-21"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
```

### Problem 1: Penguins Information and Plotting

```{r}
data("penguins", package = "palmerpenguins")
```

The variables in the `penguins` dataset are: `r colnames(penguins)`. Each row describes an individual penguin. Some important variables include `species` with
values Adelie, Gentoo, and Chinstrap, and `island` with values Torgersen, Biscoe, and Dream. The dataset includes `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r round(mean(pull(drop_na(penguins, flipper_length_mm), flipper_length_mm)),2)` mm. 

```{r}
penguins_plot = ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species)) + geom_point()

ggsave("penguins_scatterplot.png", plot = penguins_plot)
```

### Problem 2: Creating a Data Frame

```{r}
set.seed(0)

variable_df =
  tibble(
    samp_var = rnorm(10),
    log_var = samp_var > 0,
    char_var = c("a", "b", "c", "a", "b", "c", "a", "b", "c", "a"),
    fact_var = factor(char_var)
    )
```

`samp_var` mean

```{r}
variable_df |>
  pull(samp_var) |>
  mean()
```

`log_var` mean

```{r}
variable_df |>
  pull(log_var) |>
  mean()
```

`char_var` mean

```{r}
variable_df |>
  pull(char_var) |>
  mean()
```

`fact_var` mean

```{r}
variable_df |>
  pull(fact_var) |>
  mean()
```

`log_var` returns a mean because the `TRUE` and `FALSE` values are treated numerically as either 0 or 1. `char_var` returns `NA` due to the fact that it is not numerical or logical, and stores letters as its values. `fact_var` also returns `NA` due to its non numerical or logical nature. It stores its values as numbers but doesn't treat them as numeric. 

Testing out `as.numeric`

```{r warning=FALSE, results = 'hide'}
variable_df |>
  pull(log_var) |>
  as.numeric()

variable_df |>
  pull(char_var) |>
  as.numeric()

variable_df |>
  pull(fact_var) |>
  as.numeric()
```

`char_var` cannot be converted to numerical due to the current letter nature of its values. `fact_var` needed to be converted into its underlying factor category codes from its current character categories, which is why you couldn't initially take the mean. 